---
title: "Citi Bike in Jersey City"
output: flexdashboard::flex_dashboard
runtime: shiny
---
```{css}
.datatables.html-widget.html-widget-static-bound{
  height: auto !important;
  width: 90vw !important;
}
.dataTables_scrollBody{
  height: unset !important;
}
```

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(rmarkdown)
library(knitr)
library(Hmisc)
library(data.table)
library(DT)
library(tidyr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(ggcorrplot)
library(geosphere)
library(skimr)
library(leaflet)

library(data.table)
assignInNamespace("cedta.override", c(data.table:::cedta.override,"rmarkdown"), "data.table")

opts_chunk$set(echo = FALSE, comment="", warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=55), tidy = TRUE)

#opts_knit$set(root.dir = "C:/xyz/CU/5902 Capstone - Data Science Consulting/Final Project/Scripts")

#("C:/xyz/CU/5902 Capstone - Data Science Consulting/Final Project/Scripts")

Sys.setlocale("LC_TIME", "en_US.UTF-8")
```

```{r read_data}
rides = readRDS("shiny_preprocessed_rides.rds")
station_lat_lng = readRDS("station_lat_lng.rds")
```


```{r source_functions}
#source("constants.R")
#source("functions.R")
```


```{r constants}
# Most constants have been placed in the file constants.R to unify the work across reporting and applications.

# Additional constants that are unique to this reporting engine can be added here.



ride.day = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```

```{r functions}
station_hour_weekday <- function(bike_type = "Both", data) {
  if (bike_type == "Both") {
    # If not specified, use all data
    dat <- data
  } else {
    # If specified, filter data by bike_type
    dat <- data[data$rideable_type == bike_type, ]
  }

  #dat = data[rideable_type == bike_type,]
  bikes_out = dat[, .(bikes_out = .N, 
                      start_station_latitude = first(start_lat), 
                      start_station_longitude = first(start_lng)), 
                by = .(station_name = start_station_name, date = ride_date, week_day = ride_day, hour = hour(started_at))]
  
  bikes_in = dat[, .(bikes_in = .N, 
                      end_station_latitude = first(end_lat), 
                      end_station_longitude = first(end_lng)), 
                by = .(station_name = end_station_name, date = ride_date, week_day = ride_day, hour = hour(started_at))]

  bikes_change = merge(bikes_in, bikes_out, by = c("station_name", "date", "week_day", "hour"), all = TRUE)

  bikes_change[is.na(bikes_change),] = 0
  #bikes_change = bikes_change[, net_change := bikes_in - bikes_out,]
  setorderv(bikes_change, cols = c("station_name", "date", "hour"))
  #bikes_change = bikes_change[, cumulative := cumsum(net_change), by = station_name]
  bikes_change = bikes_change[, .(station_name, date, week_day, hour, bikes_in, bikes_out,
    latitude = coalesce(end_station_latitude, start_station_latitude), 
    longitude = coalesce(end_station_longitude,start_station_longitude),
    bike_type = bike_type
    )]
}

```



Introduction
=====================================  
This is the dynamic application for our Citi Bike in Jersey City project. There are 5 tabs in this app. Users can select top routes on different weekdays and see the routes and ride numbers. In the second tab, users can see station-level bike flows. The third tab is to compare the number of rides with electric bikes and classic bikes throughout a day at different stations on different weekdays. The fourth tab shows ride counts. The fifth tab shows distributions of riding duration, distance, and speed for the two types of bikes.


Routes
=====================================  

Row {data-height=500}
-------------------------------------

```{r routes_input}
inputPanel(
  selectInput(inputId = "ride_day", label = "Select the Weekday:", choices = ride.day, selected = ride.day[1]),
  #checkboxInput(inputId = "product_info_decreasing", label = "Sorted", value=TRUE),
  sliderInput(inputId = "top_n", label = "Top N Routes", min = 1, max = 20, value = 5, step = 1)
)
```

Row {data-height=500}
-------------------------------------

```{r routes_output}
renderLeaflet({
  rides_tab = rides[ride_day %in% input$ride_day,]
  rides_tab = rides_tab[, .(count = .N,
                            start_lat = first(start_lat), 
                            start_lng = first(start_lng),
                            end_lat = first(end_lat),
                            end_lng = first(end_lng),
                            start_station_name = first(start_station_name),
                            end_station_name = first(end_station_name)), 
                       by = c("station_path", "rideable_type")]
  setorderv(x = rides_tab, cols = "count", order = -1)
  rides_tab = rides_tab[1:input$top_n,]
  
  #setnames(rates, "get", input$respondent_variable)
  
  # Map
  #leaflet(data = rides_tab) %>%
  #addTiles() %>%
  #addPolylines(lng = ~c(start_lng, end_lng), 
  #             lat = ~c(start_lat, end_lat), 
  #             weight = ~count,
  #             color = "blue", opacity = 1)
  
  
  library(RColorBrewer)

# Define the number of levels in your data (e.g., the maximum count of rides)
num_levels <- max(rides_tab$count)

# Create a color ramp function interpolating between the colors of the chosen palette
color_ramp <- colorRampPalette(brewer.pal(9, "Blues"))(num_levels)
#color_ramp <- colorRampPalette(c("lightblue", "#000033"))(num_levels)  

# Create a function to map a count to a color
get_color <- function(count, max_count) {
  # Normalize count to the range of 1 to num_levels
  normalized_count <- round((count / max_count) * num_levels)
  # Use the normalized count to get the color
  color_ramp[normalized_count]
}
  
  m <- leaflet() %>% 
  addTiles() %>%  # Add default OpenStreetMap map tiles
  setView(lng = -74.067622, lat = 40.725340, zoom = 13)

  for(i in 1:nrow(rides_tab)) {
    label <- paste("Start:", rides_tab$start_station_name[i], 
                 "<br>End:", rides_tab$end_station_name[i],
                 "<br>Rides:", rides_tab$count[i])
    
    m <- m %>% addPolylines(
      lng = c(rides_tab$start_lng[i], rides_tab$end_lng[i]),
      lat = c(rides_tab$start_lat[i], rides_tab$end_lat[i]),
      color = get_color(rides_tab$count[i], num_levels),
      weight = 4,
      opacity = 0.7,
      popup = label)
  }

  m
  
})
```


Stations
=====================================  

Row {data-height=500}
-------------------------------------

```{r read_tab2_data}
change = readRDS("stations_tab.rds")

```


```{r tab2_constants}
dates = unique(change$date)
bike_type = unique(change$bike_type)
```


```{r tab2_input}
inputPanel(
  checkboxGroupInput("metrics", "Select Metrics:",
                     choices = list("Bikes In" = "bikes_in",
                                    "Bikes Out" = "bikes_out",
                                    "Net Change" = "net_change",
                                    "Cumulative" = "cumulative")),
  selectInput("bike_type", "Select Bike Type:",
              choices = bike_type),
  selectInput("date", "Select Date:", choices = dates)
)
```

Row {data-height=500}
-------------------------------------

```{r tab2_output}
renderLeaflet({
  selected_date <- input$date
  selected_metrics <- input$metrics
  selected_bike_type <- input$bike_type
  filtered_data <- change[date == selected_date & bike_type == selected_bike_type,]


  # Create popups
  popups <- sapply(1:nrow(filtered_data), function(i) {
    row <- filtered_data[i]
    info <- paste(sapply(selected_metrics, function(metric) {
      paste0(metric, ": ", row[[metric]])
    }), collapse = "<br/>")
    paste0("<strong>Station: ", row$station_name, "</strong><br/>", info)
  })

  m <- leaflet(filtered_data) %>%
    addTiles() %>%
    setView(lng = -74.067622, lat = 40.725340, zoom = 13) %>%
    
    addMarkers(~longitude, ~latitude, popup = popups)
  
  m
})
```



E-Bikes / Classic Bikes
=====================================  

Row {data-height=500}
-------------------------------------

```{r tab3_read_data}
hw_avg = readRDS("electric_vs_classic.rds")

```

```{r tab3_constants}
week_day = unique(hw_avg$week_day)
bike_type_tab3 = c("Both","electric vs classic")
```

```{r tab3_input}
fluidRow(
column(width = 6,
       inputPanel(
  selectInput("bike_type_tab3", "Select Bike Type:",
              choices = bike_type_tab3,
              selected = bike_type_tab3[1]),
  selectInput("week_day", "Select Date:", choices = week_day)
       )),
  
  column(width = 12,
         plotOutput("plot"))
)
```

```{r tab3_map_input}
fluidRow(
  column(width = 12,
         leafletOutput("map")
  )
)
```



Row {data-height=500}
-------------------------------------

```{r tab3_output}

output$map <- renderLeaflet({
  # Your Leaflet map initialization code should go here
  # For example:
  leaflet(station_lat_lng) %>%
    addTiles() %>%
    setView(lng = -74.067622, lat = 40.725340, zoom = 13) %>%
    addMarkers(~longitude, ~latitude, layerId = ~station_name, popup = ~station_name)
  })

# Define a reactive expression that will update whenever the inputs change.
filtered_data <- reactive({
  req(input$map_marker_click)  # Require that a map marker has been clicked
  clicked_id <- input$map_marker_click$id

  week_day_val <- input$week_day
  bike_type_val <- input$bike_type_tab3

  # Print the input values and their types
  print(paste("Clicked ID:", clicked_id))
  print(paste("Week Day:", week_day_val))
  print(paste("Bike Type:", bike_type_val))

  # Check data types
  print(typeof(hw_avg$station_name))
  print(typeof(hw_avg$week_day))
  print(typeof(hw_avg$bike_type_tab3))
  
  data = hw_avg[station_name == clicked_id & 
                week_day == input$week_day & 
                group == input$bike_type_tab3, ]
})


# Render the plot based on the filtered data.
output$plot <- renderPlot({
  req(filtered_data())  
  ggplot(filtered_data(), aes(x = hour, y = avg_bikes_out, group = bike_type, color = bike_type)) +
    geom_line() +
    labs(title = paste("Data for Station", input$map_marker_click$id))
})

```


Ride Counts
=====================================  

```{r tab4_constants}
grouping.variable = c("ride_date", "ride_ym", "ride_DoW", "ride_ToD")
```


```{r tab4_input}
inputPanel(
  selectInput(inputId="grouping_variable", label = "Select Variable:", choices = grouping.variable, selected = grouping.variable[1])
)
```


Row {data-height=500}
-------------------------------------

```{r tab4_output}
renderPlot({
  data = rides[, .(count = .N), by = .(rideable_type, get(input$grouping_variable))]
  setnames(data, "get", input$grouping_variable)
  
  print(data)
  
  ggplot(data, aes(x = get(input$grouping_variable), y = count, group = rideable_type, color = rideable_type)) +
  geom_line() +
  theme_minimal() +
  labs(x = input$grouping.variable, y = "Count", title = "Ride Counts by Rideable Type and ") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
})
```



Density Plots
=====================================  

```{r tab5_constants}
dur.dist.speed = c("ride_dur_s", "ride_dist_m", "ride_speed_kmh")
member.casual = c("both", "member", "casual")
```


```{r tab5_input}
inputPanel(
  selectInput(inputId="dur_dist_speed", label = "Select Variables:", choices = dur.dist.speed, selected = dur.dist.speed[1])
)
```


Row {data-height=500}
-------------------------------------

```{r tab5_output}
renderPlot({
  data_tab5 = rides[is.na(get(input$dur_dist_speed)), input$dur_dist_speed := 0]
  #setnames(data_tab5, "get", input$grouping_variable)
  #print(data_tab5)
  
    x_limits <- switch(input$dur_dist_speed,
    "ride_dur_s" = c(0, 2000),
    "ride_dist_m" = c(0, 5000),
    "ride_speed_kmh" = c(0, 30),
    NULL  # Default case if none match
  )
    
 p = ggplot(data_tab5, aes(x = get(input$dur_dist_speed), fill = rideable_type)) +
  geom_density(alpha = 0.3)
  #scale_x_continuous(limits = c(0, 2000)) +   # This is what I changed manually
  # Add x-axis limits if they are defined
  if (!is.null(x_limits)) {
    p = p + scale_x_continuous(limits = x_limits)
  }

  p + labs(x = input$dur_dist_speed, y = 'density', fill = 'Ride Type') + 
    theme_minimal()


})
```